---
title: "AP test application"
output: pdf_document
date: "2022-08-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1 Running POST-Diff (Code adapted from Tong)

```{r Post Diff, echo = TRUE}
#' TSPostdiff AP Test
#'
#' Runs TSPostdiff in two armed scenario for n batches, with given batch size and burnin period, calculate AP Test statistic for each arm.
#' @param pa prior probability of arm a
#' @param pb prior probability of arm b
#' @param n number of trials in one simulation 
#' @param burnin number of batches that runs only uniform random sampling before applying TS-Post Diff to provide more successes and failures in the beta distribution # nolint
#' @param batch number of trials before updating posterior probability
#' 
#' @return vector containing WaldScore[n], reward[n], APT_1, APT_2: the final waldscore, reward, and AP test statistics for both arms

TSPDD <- function(pa, pb, n, c, burnin, batch) {
  # Track successes and failures under each arm
  # ex. [0,1,1,1,2,3,3,4,5,6] final index indicates total number total successes
  arm_a_successes <- c(0)
  arm_a_failures <- c(0)
  arm_b_successes <- c(0)
  arm_b_failures <- c(0)
  
  # Current success and failure (changes for each trial)
  arm_a_s <- c(0)
  arm_a_f <- c(0)
  arm_b_s <- c(0)
  arm_b_f <- c(0)

  # History that tracks if there was success in either arm for each trial
  # ex. [1,0,1,1,0,1] 1 indicates success for either arm, 0 indicates no success for either arm
  abl <- c()
 
  # AP Test statistic for arm a
  APT_ban_UR = 0 # not including batches using UR
  APT_w_UR = 0 # including batches using UR
  
  
  # BURNIN PERIOD: use Uniform Random (UR) only
  for (i in 1:burnin) {

    # Draw a and draw b are local variables assigned random values 0-1
    draw_a <- runif(1) # Probability of drawing arm a
    draw_b <- runif(1) # Probability of drawing arm b

    # CASE 1: choose arm a
    if (draw_a > draw_b) {
      # Arm a was successful
      if (runif(1) < pa) {
        abl[i] <- 1
        arm_a_s <- arm_a_s + 1 # update arm a successes
      } 
      # Arm a was not successful
      else {       
        abl[i] <- 0
        arm_a_f <- arm_a_f + 1 # update arm a failures
      }
    }

    # CASE 2: choose arm b
    else{
      # Arm b was successful
      if (runif(1) < pb) {
        abl[i] <- 1
        arm_b_s <- arm_b_s + 1 # update arm b successes
      }
      # Arm b was not successful
      else {
        abl[i] <- 0
        arm_b_f <- arm_b_f + 1 # update arm b failures
      }
    }

    # Update total number of success and failures for arms a and b
    arm_a_successes[i] <- arm_a_s
    arm_a_failures[i] <- arm_a_f
    arm_b_successes[i] <- arm_b_s
    arm_b_failures[i] <- arm_b_f 
     
    #tracking the build-up of successes and failures under each arm, during
    #each iteration of for loop: starts at 0, continues until #burnin                       
  }

  # NOT BURNIN PERIOD: Use TS or UR depending on difference between posterior probabilities
  for (i in (burnin + 1):n) {

    # Posterior probability is updated according to batch size (when m = 0)
    m = i %% batch 

    # Update posterior probability of arms a and b
    if (m == 0){
      # draw from Beta(alpha, beta) distribution based on current successes & failures 
      # alpha = current successes, beta = current failures of each arm
      draw_a <-
        rbeta(1, arm_a_successes[i - 1] + 1, arm_a_failures[i - 1] + 1) 
      draw_b <-
        rbeta(1, arm_b_successes[i - 1] + 1, arm_b_failures[i - 1] + 1) 
    }

    # Does not update posterior probability 
    else{
      # i-m indicates that we are using the posterior probabilities from the previous batch
      draw_a <-
        rbeta(1, arm_a_successes[i - m] + 1, arm_a_failures[i - m] + 1) 
      draw_b <-
        rbeta(1, arm_b_successes[i - m] + 1, arm_b_failures[i - m] + 1)
    }

    # Use UR: if difference between the posterior probabilities of the two arms is smaller than the threshold, 
    if (abs(draw_a - draw_b) < c) {
      draw_a <- runif(1)
      draw_b <- runif(1)
    } 
    # Use TS otherwise
    else {
      if (m == 0){ #greater batch size -> slower and less post prob updates
        draw_a <-
          rbeta(1, arm_a_successes[i - 1] + 1, arm_a_failures[i - 1] + 1)
        draw_b <-
          rbeta(1, arm_b_successes[i - 1] + 1, arm_b_failures[i - 1] + 1)
      }
      else{
        draw_a <-
          rbeta(1, arm_a_successes[i - m] + 1, arm_a_failures[i - m] + 1)
        draw_b <-
          rbeta(1, arm_b_successes[i - m] + 1, arm_b_failures[i - m] + 1)
      }
      
      # Update AP_{ban UR} (without UR sampling)
      if (m == 0){
        if (draw_b >draw_a){
          APT_ban_UR = APT_ban_UR + 1
        }
      }
    }

    # Update AP_{w/ UR} (including the ones with UR)
    if(m == 0){
      if (draw_b > draw_a){
        APT_w_UR <- APT_w_UR + 1
      }
    }

    # Update arm a and arm b history, successes, and failures
    if (draw_a > draw_b) {
      if (runif(1) < pa) {
        abl[i] <- 1
        arm_a_s <- arm_a_s + 1
      } else {
        abl[i] <- 0
        arm_a_f <- arm_a_f + 1
      }
    } else{
      if (runif(1) < pb) {
        abl[i] <- 1
        arm_b_s <- arm_b_s + 1
      } else {
        abl[i] <- 0
        arm_b_f <- arm_b_f + 1
      }
    }  
    arm_a_successes[i] <- arm_a_s
    arm_a_failures[i] <- arm_a_f
    arm_b_successes[i] <- arm_b_s
    arm_b_failures[i] <- arm_b_f
  }


  # (list) running history of number of times arm A and B were pulled
  na <- arm_a_successes + arm_a_failures
  nb <- arm_b_successes + arm_b_failures

  # (list) running history of posterior prob for arm A and B to be successful
  pa_est <- arm_a_successes / na
  pb_est <- arm_b_successes / nb

  # Wald Score for confidence interval: is a way to analyze categorical data underlying chi-squared distribution vectors, instead of summation
  WaldScore <-
    (pa_est - pb_est) / sqrt(pa_est * (1 - pa_est) / na + pb_est * (1 - pb_est) /nb)
  
  # (list) expected reward we have from both arms at different sample sizes
  reward <- (na * pa + nb * pb) / c(1:n)
  
  return(c(WaldScore[n], reward[n], APT_ban_UR, APT_w_UR))
}
```


## Finding Critical Values using Normal Assumption on Law of Large Numbers:

```{r}
#' Critical Value (Normal)
#'
#' Finds critical value using normal assumption on LLN
#' @param APT AP Test statistic (list)
#' @param alpha Customized significance level (float)
#' 
#' @return generated critical value from normality assumption

normal_cv <- function(APT, alpha){
  m <- mean(APT) 
  v <- var(APT)
  sd <- sqrt(v)
  normal_rej <- qnorm(1-alpha)
  cv <- normal_rej * sd + m
  return(cv)
}
```

## Finding Critical Values using Empirical Assumption on 5000 simulations:

```{r}
#' Critical Value (Empirical)
#'
#' Finds critical value using empirical assumption on LLN
#' @param APT AP Test statistic (list)
#' @param alpha Customized significance level (float)
#' 
#' @return generated critical value from empirical simulations

empirical_cv <- function(APT, alpha){
  cdf = 0
  i = 0 

  # Repeat until cdf reaches 1-significance level
  while (cdf <= 1- alpha){
    p_i <-sum(APT == i)/length(APT) # Actual probability from empirical values
    cdf <- sum(cdf, p_i)
    i = i + 1
  }
  return(i-1)
}
```



## Calculating FPR for case where effect size is 0, to find critical values for APT_{ban UR}
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0 # effect size is 0: no difference between arms
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2
N <- 150 # sample_size(total number of trials in one simulation)
B <- 5000 # number of simulations
c <- 0.1 # threshold
burn_in <- 2 # # of burnin trials
n <-  3 # batch_size
Ti <- N/n # num_batches


# One output for each simulation, the four dimensions are waldscore, average reward, APT_ban_UR, and APT_w_UR
results <- array(dim = c(B, 4))

# number of simulations
for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = N,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1, burnin=2 means there is one UR distribution before apply TSPost-Diff
      batch = n
    ) 
}

average_power = mean(results[, 1] > qnorm(1-alpha)) # Find average power based on waldscore
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]

#Histogram of APT_ban_UR
hist(APT_ban_UR, freq = FALSE)

# Empirical and normal CV of APT_ban_UR is
em_cv <- empirical_cv(APT_ban_UR,alpha)
paste('empirical CV is', em_cv)
norm_cv <- normal_cv(APT_ban_UR,alpha)
paste('normal CV is', norm_cv)

# Empirical cdf of APT_ban_UR is
F1 <- ecdf(APT_ban_UR)

# Empirical and normal FPR of APT_ban_UR is
paste('empirical FPR is', 1-F1(em_cv))
paste('normal FPR is', 1-F1(floor(norm_cv)))
```

## Calculating power when p2 -p1 = 0.05
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0.05 # effect size is 0.05: difference between arm's rewards is 0.05
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2

for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = N,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    ) 
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]

#Histogram of APT_ban_UR
hist(APT_ban_UR, freq = FALSE)
F1 <- ecdf(APT_ban_UR)

paste('empircal Power is', 1-F1(em_cv))
paste('normal Power is', 1-F1(floor(norm_cv)))
```
## Calculating power when p2 -p1 = 0.1
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0.1
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2

for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = Ti,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    ) 
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]
#Histogram of APT_ban_UR
hist(APT_ban_UR, freq = FALSE)

F1 <- ecdf(APT_ban_UR)

paste('empircal Power is', 1-F1(em_cv))
paste('normal Power is', 1-F1(floor(norm_cv)))
```

## Calculating power when p2 -p1 = 0.2
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0.2
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2

for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = Ti,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    ) 
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]
#Histogram of APT_ban_UR
hist(APT_ban_UR, freq = FALSE)

F1 <- ecdf(APT_ban_UR)

paste('empircal Power is', 1-F1(em_cv))
paste('normal Power is', 1-F1(floor(norm_cv)))
```
## Calculating FPR  for case where effect size is 0, to find critical values for APT_w_UR


```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2
results <- array(dim = c(B, 4))
for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = Ti,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    ) 
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]
#Histogram of APT_w_UR
hist(APT_w_UR, freq = FALSE)

# emipircal CV of APT_w_UR is
em_cv <- empirical_cv(APT_w_UR,alpha)
paste('empirical CV is', em_cv)
# normal CV of APT_w_UR is

norm_cv <- normal_cv(APT_w_UR,alpha)
paste('normal CV is', norm_cv)

# empirical cdf of APT_w_UR is

F1 <- ecdf(APT_w_UR)

paste('empircal FPR is', 1-F1(em_cv))
paste('normal FPR is', 1-F1(floor(norm_cv)))
```

## Calculating power when p2 -p1 = 0.05
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0.05
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2

for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = Ti,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    ) 
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]
#Histogram of APT_w_UR
hist(APT_w_UR, freq = FALSE)

F1 <- ecdf(APT_w_UR)

paste('empircal Power is', 1-F1(em_cv))
paste('normal Power is', 1-F1(floor(norm_cv)))
```
## Calculating power when p2 -p1 = 0.1
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0.1
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2

for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = Ti,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    ) 
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]
#Histogram of APT_w_UR
hist(APT_w_UR, freq = FALSE)

F1 <- ecdf(APT_w_UR)

paste('empircal Power is', 1-F1(em_cv))
paste('normal Power is', 1-F1(floor(norm_cv)))
```

## Calculating power when p2 -p1 = 0.2
```{r FPR, echo=TRUE}
set.seed(0)
#calculate power and reward in case p0 < p1
alpha <- 0.05
effect_size <- 0.2
p0 <- 0.5 - effect_size / 2
p1 <- 0.5 + effect_size / 2

for (i in 1:B) {
  results[i,] <-
    TSPDD(
      pa = p0,
      pb = p1,
      n = Ti,
      c = c,
      burnin = burn_in, #burnin=1 means no burnin needs to be larger than batch-1
      batch = n
    )
    )
}
average_power = mean(results[, 1] > qnorm(1-alpha))
average_reward = mean(results[, 2])
APT_ban_UR = results[,3]
APT_w_UR = results[,4]
#Histogram of APT_w_UR
hist(APT_w_UR, freq = FALSE)

F1 <- ecdf(APT_w_UR)

paste('empircal Power is', 1-F1(em_cv))
paste('normal Power is', 1-F1(floor(norm_cv)))
```

## check for normality:
```{r}
m <- mean(APT_w_UR)
sd <- sqrt(var(APT_w_UR))

hist((APT_w_UR-m)/sd)
```

