---
title: "Nina's result replication"
author: "AP Test Team"
format: pdf
editor: visual
  warning: false
  message: false
---

# Replication of Nina's results:

This file is aimed to replicate the results from Nina's Allocation Probability Test Paper:

At: https://arxiv.org/pdf/2111.00137.pdf

Especially graphs at A4

It is to provide the correct sanity check before moving into using magnitude to compare between the results.

The .stan code for generating a K arm gaussian thompson sampling setting is at: gaussian_ts.stan

A demo of running such is below:

```{r}
library(rstan) # enable rstan, feel free to use install.packages('rstan')
library(tidyverse) # for simplicity of changing tibbles
library(tidybayes) # for simiplicity of viewing stan results
library(here) # universal access to the ap folder
```

```{r}
# let's run a quick simulation, with 4 arms, and 1000 iterations, 
# each arm has reward N(0,1)

# Define the data for the simulation
K <- 4 # number of arms
N <- 1000 # number of participants (trials)
mu0 <- c(0, 0, 0, 0) # prior means of the arms
sigma0 <- 1 # prior standard deviation of the arms
sigma <- 1 # true standard deviation of the rewards

# Compile the Stan model
stan_model <- stan_model(file = "code_check/gaussian_ts.stan")

# Run the simulation
sim_data <- list(K = K,
                 N = N,
                 mu0 = mu0, 
                 sigma0 = sigma0, 
                 sigma = sigma)
sim_results <- sampling(stan_model, 
                        data = sim_data, 
                        iter = 1000,  # number of simulations
                        chains = 4) # on how many chains
                        #algorithm = "Fixed_param")

# Extract the results
rewards <- extract(sim_results, "rewards")$rewards
chosen_arm <- extract(sim_results, "chosen_arm")$chosen_arm

sim_results

```

now let's move to a two arm setting with it's allocation probability, this one it is in gaussian_2arm.stan

```{r}

# Define the data for the simulation
N <- 1000 # number of trials
mu0 <- c(0, 0) # prior means of the arms
              # remember arm 1 is experimental, arm 2 is controlled
sigma0 <- 1 # prior standard deviation of the arms
sigma <- 1 # true standard deviation of the rewards

# Compile the Stan model
stan_model <- stan_model(file = "gaussian_2arm.stan")

# Run the simulation
sim_data <- list(N = N, 
                 mu0 = mu0, 
                 sigma0 = sigma0, 
                 sigma = sigma)
sim_results <- sampling(stan_model, 
                        data = sim_data, 
                        iter = 1000, 
                        chains = 4,
                        refresh = 0)  # what is the chains doing here???hmmm the dim doesn't match lol
                        #algorithm = "Fixed_param")

# Extract the results
rewards <- extract(sim_results, "rewards")$rewards
chosen_arm <- extract(sim_results, "chosen_arm")$chosen_arm
allocation_prob <- as.tibble(t(extract(sim_results, "allocation_prob")$allocation_prob))

```

and we can take a quick look at such:

```{r}
allocation_prob |>
  pivot_longer(V1:V2000, names_to = 'experiment', values_to = 'AP') |>
  group_by(experiment) |>
  summarise(APT = sum(AP >= 0.5)) |>
  ggplot(aes(APT)) +
  geom_histogram(fill = 'lightblue') +
  theme_bw()
```
